A Toolbox for Manipulating and Assessing Color Palettes

### 2019-12-02 pkg/colorspace/paper/Toolbox/article.Rmd
small fixup
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@521 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

At the core of the \pkg{colorspace} package are various utilities for computing with
color spaces [@color:Wiki+Colorspace], as the name of the package conveys. Thus, the package helps to map various three-dimensional
representations of color to each other [@color:Ihaka:2003]. A particularly important
mapping is the one from the perceptually-based and device-independent color model
HCL (Hue-Chroma-Luminance) to standard Red-Green-Blue (sRGB) which is the basis for color
specifications in many systems based on the corresponding hex codes
[@color:Wiki+Webcolors], e.g., in HTML but also in \proglang{R}.
For completeness further standard color models are included as well in the package.
Their connections are illustrated in Figure\ \ref{fig:color-spaces}.
Color models that are (or try to be) perceptually-based are displayed with circles
and models that are not are displayed with rectangles.

@@

* `desaturate()`: Desaturate colors by chroma removal in HCL space.
* `darken()` and `lighten()`: Algorithmically lighten or darken colors in HCL and/or HLS space.
* `max_chroma()`: Compute maximum chroma for given hue and luminance in HCL space.
* `mixcolor()`: Additively mix two colors by computing their convex combination.

@@

>>>>

At the core of the \pkg{colorspace} package are various utilities for computing with
color spaces [@color:Wiki+Colorspace], as the name of the package conveys. Thus, the package helps to map various three-dimensional
representations of color to each other [@color:Ihaka:2003]. A particularly important
mapping is the one from the perceptually-based and device-independent color model
HCL (Hue-Chroma-Luminance) to standard Red-Green-Blue (sRGB) which is the basis for color
specifications in many systems based on the corresponding hexadecimal (or simply hex) codes
[@color:Wiki+Webcolors], e.g., in HTML but also in \proglang{R}.
For completeness further standard color models are included as well in the package.
Their connections are illustrated in Figure\ \ref{fig:color-spaces}.
Color models that are (or try to be) perceptually-based are displayed with circles
and models that are not are displayed with rectangles.

@@

* `desaturate()`: Desaturate colors by chroma removal in HCL space.
* `darken()` and `lighten()`: Algorithmically lighten or darken colors in HCL and/or HLS space.
* `adjust_transparency()`: Add, remove, or modify alpha transparency of colors.
* `max_chroma()`: Compute maximum chroma for given hue and luminance in HCL space.
* `mixcolor()`: Additively mix two colors by computing their convex combination.

@@

## Adjust transparency of colors

Alpha transparency is useful for making colors semi-transparent, e.g., for
overlaying different elements in graphics. An alpha value of 0 (or 00 in hex strings)
corresponds to fully transparent and an alpha value of 1 (or FF in hex strings)
corresponds to fully opaque. If a color hex string in \proglang{R} does not provide an explicit
alpha transparency, the color is assumed to be fully opaque.

color:Wiki+Webcolors

The \code{transparency} function can be used to adjust the alpha transparency
of a set of colors. It always returns a hex color specification. This hex color
can have the alpha transparency added/removed/modified depending on the
specification of \code{alpha}: \itemize{
\item \code{alpha = NULL}: Returns a hex vector with alpha transparency only if needed.
      Thus, it keeps the alpha transparency for the colors (if any) but only if
      different from opaque.
\item \code{alpha = TRUE}: Returns a hex vector with alpha transparency
      for all colors, using opaque (FF) as the default if missing.
\item \code{alpha = FALSE}: Returns a hex vector without alpha transparency for
      all colors (even if the original colors had non-opaque alpha).
\item \code{alpha} numeric: Returns a hex vector with alpha transparency for
      all colors set to the \code{alpha} argument (recycled if necessary).
}

A character vector with hexadecimal color strings with alpha transparency
corresponding to \code{alpha} argument.

## modify transparency of a color specified by name (`"black"`), hex string
(`"#000000"`), or integer position in the palette (`1`).

```{r transparency-black}
adjust_transparency("black",   alpha = c(0, 0.5, 1))
adjust_transparency("#000000", alpha = c(0, 0.5, 1))
adjust_transparency(1,         alpha = c(0, 0.5, 1))
```

three shades of gray (in different formats:
name/opaque, hex/opaque, hex/semi-transparent)

adjust transparency: only if necessary (`alpha = NULL`), add (`alpha = TRUE`),
remove (`alpha = FALSE`), or modify (`alpha = 0.8`)

```{r transparency-black}
x <- c("gray", "#BEBEBE", "#BEBEBE80")
adjust_transparency(x, alpha = NULL)
adjust_transparency(x, alpha = TRUE)
adjust_transparency(x, alpha = FALSE)
adjust_transparency(x, alpha = 0.8)
```
